HO unification from object language to meta language

### 2024-03-20 paper/appendix.tex
update appendix

  type deref-assmt subst -> assignment -> assignment -> o.
  deref-assmt S (abs T) (abs R) :- pi x\ deref-assmt S (T x) (R x).
  deref-assmt S (val T) (val R) :- deref S T R.
\end{elpicode}
% \caption{Implementation of the \Ue predicate for \Ho}
% \label{code:ue-pred}
% \Description[code:ue-pred]{code:ue-pred}
% \end{figure*}

@@

\begin{elpicode}

\end{elpicode}
% \begin{figure*}
% \caption{Implementation of the compiler}
% \label{code:ue-pred}
% \Description[code:ue-pred]{code:ue-pred}
% \end{figure*}

>>>>

  type deref-assmt subst -> assignment -> assignment -> o.
  deref-assmt S (abs T) (abs R) :- pi x\ deref-assmt S (T x) (R x).
  deref-assmt S (val T) (val R) :- deref S T R.

@@

\begin{elpicode}
  type use-binder fo.fm -> fo.fm -> o.
  use-binder N N.
  use-binder N (fo.fapp L) :- exists (use-binder N) L.
  use-binder N (fo.flam B) :- pi x\ use-binder N (B x).

  type maybe-eta fo.fm -> fo.fm -> list fo.fm -> o.
  maybe-eta N (fo.fapp[fo.fuva _|Args]) _ :- !,
    exists (x\ maybe-eta-of [] N x) Args, !. 
  maybe-eta N (fo.flam B) L :- !, pi x\ maybe-eta N (B x) [x | L].
  maybe-eta _ (fo.fapp [fo.fcon _|Args]) L :-
    split-last-n {len L} Args First Last,
    forall1 (x\ forall1 (y\ not (use-binder x y)) First) L,
    forall2 (maybe-eta-of []) {std.rev L} Last.

  type maybe-eta-of list fo.fm -> fo.fm -> fo.fm -> o.
  maybe-eta-of _ N N :- !.
  maybe-eta-of L N (fo.fapp[fo.fuva _|Args]) :- !, 
    % Head is flex -> if N is in Args, we can have eta
    forall1 (x\ exists (maybe-eta-of [] x) Args) [N|L]. 
  maybe-eta-of L N (fo.flam B) :- !, 
    pi x\ maybe-eta-of [x | L] N (B x).
  maybe-eta-of L N (fo.fapp [N|Args]) :-
    % Head is rigid -> to be an eta we consider also L
    last-n {len L} Args R,
    forall2 (maybe-eta-of []) R {std.rev L}.

  type locale ho.tm -> o.
  type get-scope-aux ho.tm -> list ho.tm -> o.
  get-scope-aux (ho.con _) [].
  get-scope-aux (ho.uva _ L) L1 :- 
    forall2 get-scope-aux L R,
    flatten R L1.
  get-scope-aux (ho.lam B) L1 :- 
    pi x\ locale x => get-scope-aux (B x) L1.
  get-scope-aux (ho.app L) L1 :- 
    forall2 get-scope-aux L R,
    flatten R L1.
  get-scope-aux X [X] :- name X, not (locale X).
  get-scope-aux X [] :- name X, (locale X).

  type get-scope ho.tm -> list ho.tm -> o.
  get-scope T Scope :- names N,
    get-scope-aux T ScopeDuplicata,
    std.filter N (mem ScopeDuplicata) Scope.

  type close-links (ho.tm -> links) -> links -> o.
  close-links (_\[]) [].
  close-links (v\[L|XS v]) [L|YS] :- !, close-links XS YS.
  close-links (v\[(L v)|XS v]) [ho.abs L|YS] :- !, 
    close-links XS YS.

  type comp-lam (fo.fm -> fo.fm) -> (ho.tm -> ho.tm) -> 
    mappings -> mappings -> links -> links -> ho.subst -> 
      ho.subst -> o.
  comp-lam F F1 Mappings1 Mappings2 L L2 S S1 :-
    pi x y\ (pi M L S\ comp x y M M L L S S) => 
      comp (F x) (F1 y) Mappings1 Mappings2 L (L1 y) S S1, 
        close-links L1 L2.

  type comp fo.fm -> ho.tm -> mappings -> mappings -> links -> 
    links -> ho.subst -> ho.subst -> o.
  comp (fo.fcon X) (ho.con X)        Map1 Map1 L1 L1 S1 S1.
  comp (fo.flam F) (ho.uva E0 Scope) Map1 Map2 L1 L3 S1 S3 :-
    (pi x\ maybe-eta x (F x) [x]), !,
    mem.new S1 E0 S2,
    comp-lam F F1 Map1 Map2 L1 L2 S2 S3,
    (pi x\ locale x => get-scope (F1 x) Scope), 
    NewLink = ho.val(link-eta (ho.uva E0 Scope) (ho.lam F1)), 
    L3 = [NewLink| L2].
  comp (fo.flam F) (ho.lam F1)       Map1 Map2 L1 L2 S1 S2 :- 
    comp-lam F F1 Map1 Map2 L1 L2 S1 S2.
  comp (fo.fuva N) (ho.uva M []) Map1 Map2 L L S S1 :- 
    alloc Map1 Map2 (fv N) (hv M (arity z)) S S1.
  comp (fo.fapp [fo.fuva N|Ag]) (ho.uva M' Scope) Map1 Map3 L1 L3 S1 S4 :- !,
    split-pf Ag PF NPF,
    forall2 (x\y\ comp x y _ _ _ _ _ _) PF PF1,
    fold6 comp NPF NPF1 Map1 Map2 L1 L2 S1 S2,
    len PF Arity, 
    alloc Map2 Map3 (fv N) (hv M (arity Arity)) S2 S3,
    if (NPF = []) 
      (Scope = PF1, M' = M, S3 = S4, L2 = L3)
      (get-scope (ho.app {std.append PF1 NPF1}) Scope, 
        mem.new S3 M' S4,
        L3 = [@val-link-beta (ho.uva M' Scope) (ho.app[ho.uva M PF1 | NPF1]) | L2]).
  comp (fo.fapp A) (ho.app A1) Map1 Map2 L1 L2 S1 S2 :- 
    fold6 comp A A1 Map1 Map2 L1 L2 S1 S2.

  type make-eta-link-aux nat -> address -> address -> 
    list ho.tm -> links -> ho.subst -> ho.subst -> o.
  make-eta-link-aux z Ad1 Ad2 Scope1 L H1 H1  :-
    std.rev Scope1 Scope, eta-expand (ho.uva Ad2 Scope) @one T1,
    L = [@val-link-eta (ho.uva Ad1 Scope) T1].
  make-eta-link-aux (s N) Ad1 Ad2 Scope1 L H1 H3 :-
    std.rev Scope1 Scope, mem.new H1 Ad H2, 
    eta-expand (ho.uva Ad Scope) @one T2,
    (pi x\ make-eta-link-aux N Ad Ad2 [x|Scope1] (L1 x) H2 H3),
    close-links L1 L2,
    L = [@val-link-eta (ho.uva Ad1 Scope) T2 | L2].

  type make-eta-link nat -> nat -> address -> address -> 
          list ho.tm -> links -> ho.subst -> ho.subst -> o.
  make-eta-link (s N) z Ad1 Ad2 Vars L H H1 :- 
    make-eta-link-aux N Ad2 Ad1 Vars L H H1.
  make-eta-link z (s N) Ad1 Ad2 Vars L H H1 :- 
    make-eta-link-aux N Ad1 Ad2 Vars L H H1.
  make-eta-link (s N) (s M) Ad1 Ad2 Vars Links H H1 :-
    (pi x\ make-eta-link N M Ad1 Ad2 [x|Vars] (L x) H H1),
    close-links L Links.

  type deduplicate-mappings mappings -> mappings -> 
          ho.subst -> ho.subst -> links -> links -> o.
  deduplicate-mappings [] [] H H L L.
  deduplicate-mappings [(mapping (fv O) (hv M (arity LenM)) as X1) | Map1] Map2 H1 H3 L1 L3 :-
    take-list Map1 (mapping (fv O) (hv M' (arity LenM'))) _, !,
    make-eta-link LenM LenM' M M' [] New H1 H2,
    std.append New L1 L2,
    deduplicate-mappings Map1 Map2 H2 H3 L2 L3.
  deduplicate-mappings [A|As] [A|Bs] H1 H2 L1 L2 :- 
    deduplicate-mappings As Bs H1 H2 L1 L2, !.
\end{elpicode}
